A boundary-integral framework to simulate \\ the viscous erosion of a porous medium

### 2018-04-09 docs/methods/erosion_methods.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a
range of scales found in nature, from large geological
formations~\cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, down
to human-scale features and patterns~\cite{daerr2003erosion}. Though
less visible, these same forces are at work at the very smallest scales,
slowly deteriorating the individual constituents of porous media
(e.g.~solid, sand, or clay) or biological structures like plaque and
biofilms~\cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such applications, we
develop numerical methods to simulate fluid-mechanical erosion of
multiple solid bodies in a Stokes flow---the most relevant regime for groundwater and bio-fluid applications.

Recent work in the high-Reynolds number regime has highlighted the
importance of the feedback between flow and morphology that occurs
during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013,
hewett2017evolution, moore2017riemann}, and a similar feedback occurs in
the related processes of dissolution and melting~\cite{Huang2015,
kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion,
hewett2017pear, claudin2017dissolution}. In such cases, the flow alters
the shape of the immersed structures, which in turn modifies the
surrounding flow, and so on. Collectively, such processes have been
referred to as `sculpting with flow'~\cite{ristroph2018sculpting}. From
this viewpoint, the example of an eroding porous medium presents new
opportunities and challenges. For example, the medium itself is composed
of many solid constituents, so multi-body interactions, especially close
interactions, must be considered.
%Second, the long-range interactions
%facilitated by Stokes flow makes all the interactions between bodies important.

@@

Next, the computed shear stress is used to evolve the solid boundaries.
Tracking the interface of fluid-structure interaction problems with a long time horizon is particularly
difficult as the mesh can become distorted and tangled. 
Furthermore, the process of erosion naturally leads to regions of high curvature and even corners.
The interface tracking issues are addressed in two ways: first, we introduce a tangential velocity that keeps the grid equispaced with respect to arc length but that does not affect the interface dynamics; second, we introduce a curvature-penalizing term into the normal interface velocity to suppress high-frequency oscillations associated with nearly singular curvature.
Furthermore, the interface is parametrized in the so-called {\thL} coordinate system so that the curvature penalization appears as a diffusive term that can easily be handled in a numerically stable fashion with an exponential integrator.

@@

The most closely related work to ours is Mitchell \& Spagnolie \cite{mit-spa2016}, who developed methods to simulate viscous erosion in three dimensions. That work employed a traction integral equation to solve for the surrounding Stokes flow. The advantage to this formulation is that it immediately provides the shear stress required to evolve boundaries. However, the standard double-layer potential used here will allow us to compute certain quantities in the fluid bulk, such as vorticity and pressure, that will facilitate our understanding of the erosion process. Furthermore, our primary aim is to simulate the erosion of multiple bodies held fixed within a porous network, whereas Mitchell \& Spagnolie simulated a single eroding body that is mobile.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a
range of scales found in nature, from large geological
formations~\cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, down
to human-scale features and patterns~\cite{daerr2003erosion}. Though
less visible, these same forces are working at the very smallest scales,
slowly deteriorating the individual constituents of porous media
(e.g.~solid, sand, or clay) or biological structures like plaque and
biofilms~\cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such applications, we
develop numerical methods to simulate fluid-mechanical erosion of
multiple solid bodies in a Stokes flow---the most relevant regime for groundwater and the aforementioned bio-fluid applications.

Recent work in the high-Reynolds number regime has highlighted the
importance of the feedback between flow and morphology that occurs
during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013,
hewett2017evolution, moore2017riemann}; a similar feedback occurs in
the related processes of dissolution and melting~\cite{Huang2015,
kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion,
hewett2017pear, claudin2017dissolution}. In such cases, the flow alters
the shape of the immersed structures, which in turn modifies the
surrounding flow, and so on. Collectively, such processes have been
referred to as `sculpting with flow'~\cite{ristroph2018sculpting}. From
this viewpoint, the example of an eroding porous medium presents new
opportunities and challenges. For example, the medium itself is composed
of many solid constituents, so multi-body interactions, especially close
interactions, must be considered.
%Second, the long-range interactions
%facilitated by Stokes flow makes all the interactions between bodies important.

@@

Next, the computed shear stress is used to evolve the solid boundaries.
Tracking the interface of fluid-structure interaction problems with a long time horizon is particularly
difficult as the mesh can become distorted and tangled. 
Furthermore, the process of erosion naturally leads to regions of high curvature and even corners.
The interface tracking issues are addressed in two ways: first, we introduce a tangential velocity that keeps the grid equispaced with respect to arc length but that does not affect the interface dynamics; second, we introduce a curvature-penalizing term into the normal interface velocity to suppress high-frequency oscillations associated with nearly singular curvature.
The interface is parametrized in the so-called {\thL} coordinate system so that the curvature penalization appears as a diffusive term that can easily be handled in a numerically stable fashion with an exponential integrator.

@@

In closely related work, Mitchell \& Spagnolie developed methods to simulate viscous erosion in three dimensions  \cite{mit-spa2016}. That work employed a traction integral equation to solve for the surrounding Stokes flow. The advantage to this formulation is that it immediately provides the shear stress required to evolve boundaries. However, the standard double-layer potential used here will allow us to compute certain quantities in the fluid bulk, such as vorticity and pressure, that will facilitate our understanding of the erosion process. Furthermore, our primary aim is to simulate the erosion of multiple bodies held fixed within a porous network, whereas Mitchell \& Spagnolie simulated a single eroding body that is mobile.
